--- 
title: "TronMake RNA-splicing documentation"
author:
  - Johannes Hausmann, TRON gGmbH
  - Jonas Ibn-Salem, TRON gGmbH
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
description: |
  This is the documentation of the TronMake RNA-splicing workflow.
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.kable.NA = '')
library(tidyverse)
```

# About TronMake RNA-splicing

TronMake RNA-splicing is a workflow to identify non-canonical tumor-specific splice junction from RNA-seq.
The workflow implements a sensitive alignment based splice junction detection and targeted re-quantification
of candidate transcript variants. In our bioinformatics pipeline, SnakeMake is employed as the primary 
workflow manager to orchestrate various steps. 

This pipeline aligns paired FASTQ files with STAR in 2-pass mode. It then calculates PSI and Intron jaccard values with FRASER.
Next, junctions are filtered for novelty. To exclude false-positve junction calls, we exclude alignments from problematic
regions and genes. Remaining junction candidates are then re-quantified using easyquant.


## Components

The pipeline currently contains the following tools

```{r components, echo = FALSE}
library(readxl)
knitr::kable(
    read_excel("resources/tools.xlsx") %>% 
        select(
            'Tool', 'Link', 'License'
        )
)
```

## Quick start guide

Clone the repository:

```
git clone --recursive https://gitlab.rlp.net/tron/tronmake-rna-splicing.git
cd tronmake-rna-splicing
```

Create and activate the conda environment containing snakemake:

```
conda env create -p conda_env -f environment.yaml
conda activate conda_env/
```

Run the pipeline:

```
snakemake \
    --config sample_sheet=</path/to/input.tsv> \
    --directory <output dir> \
    --software-deployment-method conda \
    --directory <output directory> \
    [--conda-prefix </path/to/conda/env/location>] \
    [--profile </path/to/profile/directory>]
```

## Authors & Acknowledgements 

The TronMake RNA-splice pipeline was originally developed by Johannes Hausmann at [TRON - Translational Oncology at the Medical Center of the Johannes Gutenberg University Mainz gGmbH (non-profit)](https://tron-mainz.de/).
The original intention of this pipeline was to have a robust option to quantify novel splicing in cancer RNA-seq to obtain potential neoantigen candidates.

Maintenance is now lead by Johannes Hausmann. 

Main developers: 

- [Johannes Hausmann](mailto:johannes.hausmann@tron-mainz.de)   

